### 2018-07-16 submission/manuscript.Rmd
update results

@@

**Results.** Predictably, we found noticeable differences in relative abundance based on high and low GC content (P-value $\leqslant$ `r round(max(filter(gc_data, bh < 0.05) %>% pull(bh)), digits = 2)`). Chimera prevalence in mock communities varied by polymerase with differences being most notable at 35 cycles (Kappa = `r round((filter(error_summary, cycles == "35x", taq == "K") %>% pull(chimera_prevalence_median))*100, digits = 2)`% (median) versus Platinum = `r round((filter(error_summary, cycles == "35x", taq == "PL") %>% pull(chimera_prevalence_median))*100, digits = 2)`%) and this variation persisted after chimera removal using VSEARCH. We also observed positive correlations between chimera prevalence and the number of OTUs with Platinum having the highest (R^2^ = `r max(pull(r2_table, r2))`) and Kappa having the worst (R^2^ = `r min(pull(r2_table, r2))`). When analyzing mock community samples the variation in the number of OTUs detected by the polymerases was observable as early as 20 cycles (P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`). There also was a large range in the number of OTUs amplified by the polymerases at 35 cycles (Accuprime = `r filter(mock_num_otu_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_min)` - `r filter(mock_num_otu_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_max)` versus Phusion = `r filter(mock_num_otu_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(mock_num_otu_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`). When analyzing fecal samples we observed similar differences in the number of OTUs. However, the median number of OTUs only varied by HiFi DNA polymerase used at 35 cycles and not at 20 cycles like the mock community (P-value `r ifelse(max(filter(fs_anova_summary, cycle == "35x")[, "bh"]) < 0.0001, invisible("<"), invisible("check data"))` 0.0001). Random Forest models were most successful at classifying individual donor samples rather than polymerase or number of cycles used (P-value $\leqslant$ `r format(max(filter(dunn_rf_data, str_detect(var1, "fecal"), str_detect(var2, "fecal")) %>% pull(bh)), scientific = T, digits = 3)`). Additionally, the most important OTUs in the polymerase and number of cycle models were not the most important in the individual donor sample model.

**Conclusions.** Although there are 16S rRNA gene sequencing differences based on polymerase and number of cycles used, they are small with respect to the biological differences between fecal bacterial communities within individuals. Collectively, these results provide evidence that a real biological difference between groups, based on 16S rRNA gene sequencing of fecal samples, should be detectable regardless of polymerase and number of cycles used.

@@

We next investigated the effect polymerases and number of cycles had on the bacterial community composition in more detail by using the Bray-Curtis index. We observed no difference in the bacterial community composition between any measure tested when using the mock community (P-value `r ifelse(pull(mock_kruskal_test_by_cycle_compare_summary, p.value) < 0.05, invisible("<"), invisible(">"))` 0.05) [Figure 3B]. Finding minimal differences in overall bacterial community composition, we next examined if specific taxa vary depending on polymerase and number of cycles. We found that OTUs that taxonomically classified to *Salmonella*, *Escherichia*, *Enterococcus*, and *Staphylococcus* differed based on polymerase used [Table S6 & S7]. These specific OTUs were all members of the mock community and the differences were most pronounced at 25 and 30 cycles [Table S7]. The majority of the differences between taxa were due to differences between a single polymerase (Kappa) and the other polymerases tested [Table S7].

@@

When using the Bray-Curtis index, we observed that the affect polymerases and number of cycles had on the bacterial community composition in fecal samples was minimal. Within each respective 5-cycle increment comparison there was no difference in Bray-Curtis index between the polymerases (P-value `r ifelse(min(filter(bray_sim_to_prev_cycle_kruskal_results, sub_sample_level == 1000)[, "bh"], na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05) [Figure 4B]. When using PERMANOVA to test for community differences based on any of the number of cycles within polymerases, only Phusion had cycle dependent differences (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`). Additionally, Phusion was one of two polymerases that had enough sequences to be rarefied to 1000 sequences at 15 cycles. There was also an overall decrease in Bray-Curtis index by cycle comparison group (e.g. the 15 cycle versus 20 cycle group compared to the 30 cycle versus 35 cycle group) (P-value `r ifelse(pull(fecal_kruskal_test_by_cycle_compare_summary, p.value) < 0.01, invisible("<"), invisible(">"))` 0.01) [Figure 4B]. Using a Dunn's post-hoc test the 15 cycle versus 20 cycle and 20 cycle versus 25 cycle comparison groups had a higher Bray-Curtis index than the 30 cycle versus 35 cycle comparison group (P-value `r ifelse(min(pull(dunns_test_by_cycle_compare_summary, P.adj)) < 0.05, invisible("<"), invisible(">"))` 0.05). In addition to assessing community composition by Bray-Curtis index, we also investigated differences in OTUs based on polymerase at each number of cycle used and no significant differences could be found (P-value $\geqslant$ `r round(min(pull(fecal_otu_tests, p.value)), digits = 2)`, corrected P-value = `r round(min(pull(fecal_otu_tests, bh)), digits = 2)`). Overall, these results suggest that in fecal samples the number of cycles and polymerase used only minimally change the bacterial community composition.

>>>>

bray_mock_combined <- read_csv("../data/process/tables/mock_bray_combined_permanova_results.csv")
bray_fecal_combined <- read_csv("../data/process/tables/fecal_bray_combined_permanova_results.csv")

@@

**Results.** Predictably, we found noticeable differences in relative abundance based on high and low GC content (P-value $\leqslant$ `r round(max(filter(gc_data, bh < 0.05) %>% pull(bh)), digits = 2)`). Chimera prevalence in mock communities varied by polymerase with differences being most notable at 35 cycles (Kappa = `r round((filter(error_summary, cycles == "35x", taq == "K") %>% pull(chimera_prevalence_median))*100, digits = 2)`% (median) versus Platinum = `r round((filter(error_summary, cycles == "35x", taq == "PL") %>% pull(chimera_prevalence_median))*100, digits = 2)`%) and this variation persisted after chimera removal using VSEARCH. We also observed positive correlations between chimera prevalence and the number of OTUs with Platinum having the highest (R^2^ = `r max(pull(r2_table, r2))`) and Kappa having the worst (R^2^ = `r min(pull(r2_table, r2))`). When analyzing mock community samples the variation in the number of OTUs detected by the polymerases was observable as early as 20 cycles (P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`). There also was a large range in the number of OTUs amplified by the polymerases at 35 cycles (Accuprime = `r filter(mock_num_otu_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_min)` - `r filter(mock_num_otu_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_max)` versus Phusion = `r filter(mock_num_otu_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(mock_num_otu_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`). When analyzing fecal samples we observed smaller differences in the number of OTUs. Additionally, the median number of OTUs only varied by HiFi DNA polymerase used at 35 cycles and not at 20 cycles like the mock community (P-value `r ifelse(max(filter(fs_anova_summary, cycle == "35x")[, "bh"]) < 0.0001, invisible("<"), invisible("check data"))` 0.0001). Random Forest models were most successful at classifying individual donor samples rather than polymerase or number of cycles used (P-value $\leqslant$ `r format(max(filter(dunn_rf_data, str_detect(var1, "fecal"), str_detect(var2, "fecal")) %>% pull(bh)), scientific = T, digits = 3)`). Additionally, the most important OTUs in the polymerase and number of cycle models were not the most important in the individual donor sample model.

**Conclusions.** Although there were 16S rRNA gene sequencing differences based on polymerase and number of cycles used, bacterial community composition differences are mostly only detectable in mock communities. Collectively, these results provide evidence that smaller biological differences between groups, based on 16S rRNA gene sequencing of fecal samples, can be consistently detected regardless of polymerase and number of cycles used.

@@

We next investigated the effect polymerases and number of cycles had on the bacterial community composition in more detail by using the Bray-Curtis index. We observed clear differences in the mock community based on polymerase used (P-value = `r format(filter(bray_mock_combined, x1 == "polymerase") %>% select("Pr(>F)") %>% pull(), scientific = T, digits = 2)`) as well as differences based on polymerase and number of cycles (P-value = `r format(filter(bray_mock_combined, x1 == "interaction_polymerase_cycles") %>% select("Pr(>F)") %>% pull(), scientific = T, digits = 2)`). However, when investigating specific cycles within each polymerase we observed no difference in the bacterial community composition (P-value `r ifelse(pull(mock_kruskal_test_by_cycle_compare_summary, p.value) < 0.05, invisible("<"), invisible(">"))` 0.05) [Figure 3B]. Finding specific differences in overall bacterial community composition based on number of cycles and polymerase, we next examined if specific taxa vary depending on polymerase and number of cycles. We found that OTUs that taxonomically classified to *Salmonella*, *Escherichia*, *Enterococcus*, and *Staphylococcus* differed based on polymerase used [Table S6 & S7]. These specific OTUs were all members of the mock community and the differences were most pronounced at 25 and 30 cycles [Table S7]. The majority of the differences between taxa were due to differences between a single polymerase (Kappa) and the other polymerases tested [Table S7].

@@

When using the Bray-Curtis index, we observed that polymerases and number of cycles had no affect on the bacterial community composition in fecal samples (P-value = `r round(filter(bray_fecal_combined, x1 == "polymerase") %>% select("Pr(>F)") %>% pull(), digits = 2)`). There was also no interaction between individuals, number of cycles, and polymerase used (P-value = `r round(filter(bray_fecal_combined, x1 == "interaction_ind_poly_cycle") %>% select("Pr(>F)") %>% pull(), digits = 2)`). When using PERMANOVA to test for community differences based on any of the number of cycles within polymerases, only Phusion had cycle dependent differences (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`). Additionally, Phusion was one of two polymerases that had enough sequences to be rarefied to 1000 sequences at 15 cycles. Within each respective 5-cycle increment comparison there was no difference in Bray-Curtis index between the polymerases (P-value `r ifelse(min(filter(bray_sim_to_prev_cycle_kruskal_results, sub_sample_level == 1000)[, "bh"], na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05) [Figure 4B]. However, there was an overall decrease in Bray-Curtis index by cycle comparison group (e.g. the 15 cycle versus 20 cycle group compared to the 30 cycle versus 35 cycle group) (P-value `r ifelse(pull(fecal_kruskal_test_by_cycle_compare_summary, p.value) < 0.01, invisible("<"), invisible(">"))` 0.01) [Figure 4B]. Using a Dunn's post-hoc test the 15 cycle versus 20 cycle and 20 cycle versus 25 cycle comparison groups had a higher Bray-Curtis index than the 30 cycle versus 35 cycle comparison group (P-value `r ifelse(min(pull(dunns_test_by_cycle_compare_summary, P.adj)) < 0.05, invisible("<"), invisible(">"))` 0.05). In addition to assessing community composition by Bray-Curtis index, we also investigated differences in OTUs based on polymerase at each number of cycle used and no significant differences could be found (P-value $\geqslant$ `r round(min(pull(fecal_otu_tests, p.value)), digits = 2)`, corrected P-value = `r round(min(pull(fecal_otu_tests, bh)), digits = 2)`). Overall, these results suggest that in fecal samples the number of cycles and polymerase used only minimally change the bacterial community composition.

### 2018-07-16 submission/manuscript.tex
update results

\textbf{Results.} Predictably, we found noticeable differences in
relative abundance based on high and low GC content (P-value
\(\leqslant\) 0.04). Chimera prevalence in mock communities varied by
polymerase with differences being most notable at 35 cycles (Kappa =
5.71\% (median) versus Platinum = 26.62\%) and this variation persisted
after chimera removal using VSEARCH. We also observed positive
correlations between chimera prevalence and the number of OTUs with
Platinum having the highest (R\textsuperscript{2} = 0.974) and Kappa
having the worst (R\textsuperscript{2} = 0.478). When analyzing mock
community samples the variation in the number of OTUs detected by the
polymerases was observable as early as 20 cycles (P-value = 0.002).
There also was a large range in the number of OTUs amplified by the
polymerases at 35 cycles (Accuprime = 15 - 20 versus Phusion = 14 - 73).
When analyzing fecal samples we observed similar differences in the
number of OTUs. However, the median number of OTUs only varied by HiFi
DNA polymerase used at 35 cycles and not at 20 cycles like the mock
community (P-value \textless{} 0.0001). Random Forest models were most
successful at classifying individual donor samples rather than
polymerase or number of cycles used (P-value \(\leqslant\) 5.49e-07).
Additionally, the most important OTUs in the polymerase and number of
cycle models were not the most important in the individual donor sample
model.

\textbf{Conclusions.} Although there are 16S rRNA gene sequencing
differences based on polymerase and number of cycles used, they are
small with respect to the biological differences between fecal bacterial
communities within individuals. Collectively, these results provide
evidence that a real biological difference between groups, based on 16S
rRNA gene sequencing of fecal samples, should be detectable regardless
of polymerase and number of cycles used.

@@

We next investigated the effect polymerases and number of cycles had on
the bacterial community composition in more detail by using the
Bray-Curtis index. We observed no difference in the bacterial community
composition between any measure tested when using the mock community
(P-value \textgreater{} 0.05) {[}Figure 3B{]}. Finding minimal
differences in overall bacterial community composition, we next examined
if specific taxa vary depending on polymerase and number of cycles. We
found that OTUs that taxonomically classified to \emph{Salmonella},
\emph{Escherichia}, \emph{Enterococcus}, and \emph{Staphylococcus}
differed based on polymerase used {[}Table S6 \& S7{]}. These specific
OTUs were all members of the mock community and the differences were
most pronounced at 25 and 30 cycles {[}Table S7{]}. The majority of the
differences between taxa were due to differences between a single
polymerase (Kappa) and the other polymerases tested {[}Table S7{]}.

@@

When using the Bray-Curtis index, we observed that the affect
polymerases and number of cycles had on the bacterial community
composition in fecal samples was minimal. Within each respective 5-cycle
increment comparison there was no difference in Bray-Curtis index
between the polymerases (P-value \textgreater{} 0.05) {[}Figure 4B{]}.
When using PERMANOVA to test for community differences based on any of
the number of cycles within polymerases, only Phusion had cycle
dependent differences (P-value = 0.03). Additionally, Phusion was one of
two polymerases that had enough sequences to be rarefied to 1000
sequences at 15 cycles. There was also an overall decrease in
Bray-Curtis index by cycle comparison group (e.g.~the 15 cycle versus 20
cycle group compared to the 30 cycle versus 35 cycle group) (P-value
\textless{} 0.01) {[}Figure 4B{]}. Using a Dunn's post-hoc test the 15
cycle versus 20 cycle and 20 cycle versus 25 cycle comparison groups had
a higher Bray-Curtis index than the 30 cycle versus 35 cycle comparison
group (P-value \textless{} 0.05). In addition to assessing community
composition by Bray-Curtis index, we also investigated differences in
OTUs based on polymerase at each number of cycle used and no significant
differences could be found (P-value \(\geqslant\) 0.01, corrected
P-value = 1). Overall, these results suggest that in fecal samples the
number of cycles and polymerase used only minimally change the bacterial
community composition.

>>>>

\textbf{Results.} Predictably, we found noticeable differences in
relative abundance based on high and low GC content (P-value
\(\leqslant\) 0.04). Chimera prevalence in mock communities varied by
polymerase with differences being most notable at 35 cycles (Kappa =
5.71\% (median) versus Platinum = 26.62\%) and this variation persisted
after chimera removal using VSEARCH. We also observed positive
correlations between chimera prevalence and the number of OTUs with
Platinum having the highest (R\textsuperscript{2} = 0.974) and Kappa
having the worst (R\textsuperscript{2} = 0.478). When analyzing mock
community samples the variation in the number of OTUs detected by the
polymerases was observable as early as 20 cycles (P-value = 0.002).
There also was a large range in the number of OTUs amplified by the
polymerases at 35 cycles (Accuprime = 15 - 20 versus Phusion = 14 - 73).
When analyzing fecal samples we observed smaller differences in the
number of OTUs. Additionally, the median number of OTUs only varied by
HiFi DNA polymerase used at 35 cycles and not at 20 cycles like the mock
community (P-value \textless{} 0.0001). Random Forest models were most
successful at classifying individual donor samples rather than
polymerase or number of cycles used (P-value \(\leqslant\) 5.49e-07).
Additionally, the most important OTUs in the polymerase and number of
cycle models were not the most important in the individual donor sample
model.

\textbf{Conclusions.} Although there were 16S rRNA gene sequencing
differences based on polymerase and number of cycles used, bacterial
community composition differences are mostly only detectable in mock
communities. Collectively, these results provide evidence that smaller
biological differences between groups, based on 16S rRNA gene sequencing
of fecal samples, can be consistently detected regardless of polymerase
and number of cycles used.

@@

We next investigated the effect polymerases and number of cycles had on
the bacterial community composition in more detail by using the
Bray-Curtis index. We observed clear differences in the mock community
based on polymerase used (P-value = 1e-04) as well as differences based
on polymerase and number of cycles (P-value = 1e-04). However, when
investigating specific cycles within each polymerase we observed no
difference in the bacterial community composition (P-value
\textgreater{} 0.05) {[}Figure 3B{]}. Finding specific differences in
overall bacterial community composition based on number of cycles and
polymerase, we next examined if specific taxa vary depending on
polymerase and number of cycles. We found that OTUs that taxonomically
classified to \emph{Salmonella}, \emph{Escherichia},
\emph{Enterococcus}, and \emph{Staphylococcus} differed based on
polymerase used {[}Table S6 \& S7{]}. These specific OTUs were all
members of the mock community and the differences were most pronounced
at 25 and 30 cycles {[}Table S7{]}. The majority of the differences
between taxa were due to differences between a single polymerase (Kappa)
and the other polymerases tested {[}Table S7{]}.

@@

When using the Bray-Curtis index, we observed that polymerases and
number of cycles had no affect on the bacterial community composition in
fecal samples (P-value = 1). There was also no interaction between
individuals, number of cycles, and polymerase used (P-value = 1). When
using PERMANOVA to test for community differences based on any of the
number of cycles within polymerases, only Phusion had cycle dependent
differences (P-value = 0.03). Additionally, Phusion was one of two
polymerases that had enough sequences to be rarefied to 1000 sequences
at 15 cycles. Within each respective 5-cycle increment comparison there
was no difference in Bray-Curtis index between the polymerases (P-value
\textgreater{} 0.05) {[}Figure 4B{]}. However, there was an overall
decrease in Bray-Curtis index by cycle comparison group (e.g.~the 15
cycle versus 20 cycle group compared to the 30 cycle versus 35 cycle
group) (P-value \textless{} 0.01) {[}Figure 4B{]}. Using a Dunn's
post-hoc test the 15 cycle versus 20 cycle and 20 cycle versus 25 cycle
comparison groups had a higher Bray-Curtis index than the 30 cycle
versus 35 cycle comparison group (P-value \textless{} 0.05). In addition
to assessing community composition by Bray-Curtis index, we also
investigated differences in OTUs based on polymerase at each number of
cycle used and no significant differences could be found (P-value
\(\geqslant\) 0.01, corrected P-value = 1). Overall, these results
suggest that in fecal samples the number of cycles and polymerase used
only minimally change the bacterial community composition.
